---
title: "HW05_THE_FINISHER"
author: "Chad Fibke"
date: '2017-10-11'
output: github_document
---

# 1.Installation of data set and data analysis packages 

```{r}
library(gapminder)
library(tidyverse)
library(forcats)

```
Just some quick notes:

+ Comments about my code will bewithin code chunks:
```{r}
# Like this 
```
+ Comments about my workflow will be outside of the code chunks. 
+ I will be working with the gapminder dataset!

# 2.Factor management 
Let's take a look at gapminder to see what variables are concidered factors:
```{r}
str(gapminder)
```

Now we know gapminder's variable:

+ Country is a factor and  has 142 different levels.

```{r Country.levels, echo=TRUE}
levels(gapminder$country) %>% 
  head() %>% 
  knitr::kable(col.names = "Country Factors")
  # I'm only showing few levels to help our poor eyes
```

+ Continent is a factor and has 5 different levels. 
```{r Continent.levels, echo=TRUE}
levels(gapminder$continent) %>% 
  knitr::kable( col.names = "Continent Factors")

```

We now know what factors we are working with, so lets go ahead and see what we have to drop!

```{r}
 EX.Ocieania <- gapminder %>% 
  filter(continent == "Oceania") 
 

EX.Ocieania$country %>%   
   fct_drop() %>% 
  levels() 

EX.Ocieania$continent %>%   
   fct_drop() %>% 
  nlevels() 
```
The above shows that when we drop the Ocieania Continent: 

+ One level will be dropped for continents.
+ Two levels  will be dropped for countries.

Now we know how many we are dropping lets go in the revese direction and drop them:
```{r}
All.without.Ocieania <- gapminder %>% 
  filter(continent != "Oceania")

All.without.Ocieania$country %>%   
   fct_drop() %>% 
  nlevels()
# As seen above we have ofically droped the "Australia"   "New Zealand" levels because we now only have 140 country levels.

All.without.Ocieania$continent %>%   
   fct_drop() %>% 
  nlevels()
# As seen above we have ofically droped the Ocieania levels because we now only have 4 continent levels.

All.without.Ocieania %>% 
  filter(continent == "Ocieania") #Also shows we have lost the Ocieania continent, 

```




